Making Soup: Preparing and Validating Models of the Bacterial Cytoplasm for Molecular Simulation

### 2019-11-20 paper/rebuttal.txt

ANSWER: All fixed, thanks for finding these.

Reviewer: 2

Comments: Bortot et al. studied protein crowded systems using
molecular simulations with explicit solvent. MD simulations of
macromolecules in crowded conditions help us to understand protein
behaviors in a living cell. Recently, several computational scientists
performed MD simulations of multiple proteins in crowded conditions
and examined the differences of protein structure and dynamics in
crowded condition compared with that in dilute solution. In spite of
the importance of the research field, the paper has three critical
issues. At first, they used only small number of proteins and
tRNA(Phe) in the simulation system, although they call the set up as
the E. coli cytoplasm. Since the proteins can cover only 50 % of the
fraction of E. coli cytoplasm, it is difficult to call the system as
the E. coli cytoplasm. It should be called as a crowded condition that
mimic 50% of volume fraction of the cytosol of E. coli.

ANSWER: Obviously our model is not complete, to reflect this we
changed the title of the paper to "Making Soup: Preparing and
Validating Models of the Bacterial Cytoplasm for Molecular
Simulation". In the abstract the following sentence is added "The
model consists of 8 different proteins and one type of t-RNA molecule,
representing about 50% of the cytoplasmic proteins."

@@

They did not include ribosomes in the system. However, the size
differences between ribosomes and other proteins should be essential
in the cellular crowding effect.

ANSWER: This was a deliberate omission. A sentence was added to
explain it: "It has been estimated that a {\em E. coli} cell contains
about 50,000 ribosomes per $\mu^3$~\cite{Milo2017a}. The volume of our
box is about 12,000 nm$^3$ (\figref{shrink}), therefore in principle
one ribosome molecule should be added, however to reduce the
complexity this was omitted."

@@

Secondly, they observed tRNA
aggregation with Mg2+, ATP, and FBP in their systems. They concluded
that completely protonating the phosphate groups of ATP and FBP is
necessary to prevent their aggregation with Mg2+. However, there is no
rationale or experimental evidences on the protonation state of
ATP. Presumably, the Mg2+ model or phosphate model may have
problems. Another possibility is that they used Amber99sb-ws, where
the 1.10 scaling is used for LJ interactions between water and
macromolecules. If they applied the scaling also to tRNA, tRNA
structure could be unfolded. However, there is no data about tRNA
structural changes in crowded condition compared to that in
solution.

ANSWER: There may be a number of possible solutions to the issue. The
work by Leotyev et al. (Refs 78 and 79) offers a plausible explanation
of the problem, namely that in fixed-charge models, that is
non-polarizable models, the infinite frequency dielectric constant
should be incorporated, either by introducing a relative dielectric
constant eps(inf) of approximately 1.7, or by scaling the charges down
by a factor of sqrt(1.7). Leontyev and co-workers advocate scaling the
charges since eps(inf) is somewhat material dependent. Such a solution
was described in the manuscript already, but a new sentence was added:
"Due to the omission of the electronic polarization, the screening is
underestimated and this corresponds to applying an effective
dielectric constant of about 1.7, alternatively scale the charges down
by $\sqrt{1.7}$ which would mean that FBP would get a charge of
approximately -3 instead of -4."

@@

Thirdly, except for the protonation state, there are few novel
biophysical insights in the current work. Essentially same data on
diffusion or RMSD were already shown in the previous crowding papers,
although the length of simulations and the force fields are
different. Although the modeling strategy (Figure 1) seems to be
interesting, the rest of scientific results and discussion may have
rooms for improvements. In particular, the issue of the aggregation of
tRNA in the simulations should be overcome by the refinement of ion
models or the tRNA-water interaction scaling factors.

ANSWER: We agree that the amount of new biophysical insights is
limited. However, the purpose of this paper was rather to pave the way
for further in-depth studies by collecting information and scripts
allowing others to build on this work. For instance, as suggested by
this referee, one could add ribosomes in larger models for the E. coli
cytoplasm. We therefore think the present paper is particularly
suitable for the journal of chemical information and modeling.

### 2019-11-20 paper/virtual_cytoplasm_v2.tex
Answered referee 2

\documentclass[journal=jcisd8,manuscript=article]{achemso}
% \linespread{1.5}
\usepackage{amssymb,amsmath,graphics,epsfig}
% \usepackage{amsmath}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{float}
\usepackage{upgreek}
\usepackage{soul}
\setkeys{acs}{super = true}
\setcitestyle{super,open={},close={}}
\def\citenumfont{}

@@

\title{Making Soup: Preparing and Validating Molecular Simulations of the Bacterial Cytoplasm}

@@

\author{Zahedeh Bashardanesh}
\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
\altaffiliation{These authors contributed equally to this work}

\author{David van der Spoel}
\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology. Uppsala University, SE-751 05 Uppsala, Sweden}
\email{david.vanderspoel@icm.uu.se}

@@

\newpage
\begin{abstract}
Biomolecular crowding affects the biophysical and biochemical behavior
of macromolecules when compared to the dilute environment present in
experiments made with isolated proteins. Computational modeling and
simulation are useful tools to study how crowding affects the
structural dynamics and biological properties of macromolecules. As
computational power increased, modeling and simulating large scale
all-atom explicit solvent models of the prokaryote cytoplasm become
possible. In this work, we build an atomistic model of the cytoplasm
of \textit{Escherichia coli} composed of 1.5 million atoms and submit
it to a total of \SI{3}{\micro\second} of molecular dynamics
simulations. The properties of biomolecules under crowding conditions
are compared to those from simulations of the individual compounds
under dilute conditions.  The simulation model is found to be
consistent with experimental data about the diffusion coefficient and
stability of macromolecules under crowded conditions. In order to
stimulate further work we provide a Python script and a set of files
that enables other researchers to build their own \textit{E. coli}
cytoplasm models to address questions related to crowding.
\end{abstract}

@@

\subsection{Cytoplasm model}
In this section we describe the rationale behind the composition of
our model, which consists of five fractions: protein, RNA, metabolites, water
and ions. It should be noted that we did not add lipids and DNA to our
model because we consider only elements that are free to diffuse
through the cytosol. We gathered data from several sources in order to
build a computational model that is representative of the cytoplasm of
{\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{shrinking.pdf} 
\caption{A) The initial simulation box with $L=35.0$ nm and the box
  after the shrinking step, $L=22.9$ nm. Biomolecules in green, water in red. B) The evolution of the
  volume of the simulation box during the shrinking simulations. Each
  replica is shown in different colors. }
\end{figure}

@@

Magnesium cations were added to our cytoplasm model as counter-ions
for tRNA and ATP. However, as the simulation progressed, FBP and ATP
aggregated around Mg$^{2+}$ and, by consequence, around tRNA
% {\bf,
%   effectively removing the counter ions from the RNA. IS THAT
%   CORRECT?} {\color{blue} No, Figure \ref{fig:tRNA_aggregation} shows that Mg$^{2+}$ is still concentrated around tRNA in the soup }. 
  After identifying this problem, we used simulations of
small systems containing only the metabolites, Mg$^{2+}$ and water to
test different strategies to prevent aggregation in future studies
using long simulations of cytoplasm models. We found that protonating
the phosphate-containing metabolites is enough to reach this goal
(Fig.~\ref{fig:avoiding_aggregation}). Even though ATPH$_3$ and
FBPH$_4$ are not the biologically relevant forms of these molecules in
physiological pH, using their protonated forms may be an acceptable
compromise between parametrization complexity and practical
results. The biggest impact of this approach is on long range
interactions of these metabolites with other molecules due to the
neutralization of their net charge. However, molecules in a crowded
environment are restrained to smaller volumes and shorter inter
molecular distances, which mitigates the effect of charge
neutralization. It has been suggested that charged species should
employ scaled charges in order to be compatible with the empirical
water models~\cite{Leontyev2009a,Leontyev2011a}. Due to the high local
density of negative charges in FBP and ATP this might indeed be an
alternative solution, barring use of polarizable
models~\cite{Ponder2010a,Lopes2013a,Ghahremanpour2018b,Walz2018a}.

@@

We submitted this model to a total of \SI{3}{\micro\second} of
molecular dynamics simulations using a modified AMBER FF99SB force
field that was tuned to reproduce experimental data of systems
composed of several macromolecules~\cite{Best2014a}. The model was
validated by comparing to experimental data about the reduction of the
translational diffusion coefficient of proteins in the cytoplasm of
\textit{E. coli} when compared to the dilute condition (Figure
\ref{fig:translational_diffusion})~\cite{Elowitz1999,Konopka2006}. Additionally,
our data is in agreement with recent simulation work and experimental data
that show that the unspecific interactions between macromolecules in a
crowded environment can destabilize
them~\cite{Feig2011,miklos2011,Wang2012b}. Specifically, we can see
that, despite the overall stability of our cytoplasm model even in the
\SI{}{\micro\second} timescale, the proteins are slightly less stable
in the cytoplasm model than in dilute condition (Figures
\ref{fig:structural_integrity_chain} and \ref{fig:rmsf}).

>>>>

\documentclass[journal=jpcisd8,manuscript=article]{achemso}
% \linespread{1.5}
\setkeys{acs}{etalmode=truncate,maxauthors=0}
\usepackage{amssymb,amsmath,graphics,epsfig}
% \usepackage{amsmath}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{siunitx}
\usepackage{float}
\usepackage{upgreek}
\usepackage{soul}
\setkeys{acs}{super = true}
\setcitestyle{super,open={},close={}}
\def\citenumfont{}

@@

\title{Making Soup: Preparing and Validating Models of the Bacterial Cytoplasm for Molecular Simulation}

@@

\author{Zahedeh Bashardanesh}
\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology, Uppsala University, Box 596, SE-75124 Uppsala, Sweden}
\altaffiliation{These authors contributed equally to this work}

\author{David van der Spoel}
\affiliation{Science for Life Laboratory, Department of Cell and Molecular Biology, Uppsala University, Box 596, SE-75124 Uppsala, Sweden}
\email{david.vanderspoel@icm.uu.se}

@@

\newpage
\begin{abstract}
Biomolecular crowding affects the biophysical and biochemical behavior
of macromolecules when compared to the dilute environment present in
experiments made with isolated proteins. Computational modeling and
simulation are useful tools to study how crowding affects the
structural dynamics and biological properties of macromolecules. As
computational power increased, modeling and simulating large scale
all-atom explicit solvent models of the prokaryote cytoplasm become
possible. In this work, we build an atomistic model of the cytoplasm
of \textit{Escherichia coli} composed of 1.5 million atoms and submit
it to a total of \SI{3}{\micro\second} of molecular dynamics
simulations. 
The model consists of 8 different proteins and one type of t-RNA molecule,
representing about 50\% of the cytoplasmic proteins. 
Properties of biomolecules under crowding conditions are compared to
those from simulations of the individual compounds under dilute
conditions.  The simulation model is found to be consistent with
experimental data about the diffusion coefficient and stability of
macromolecules under crowded conditions. In order to stimulate further
work we provide a Python script and a set of files that enables other
researchers to build their own \textit{E. coli} cytoplasm models to
address questions related to crowding.
\end{abstract}

@@

\subsection{Cytoplasm model}
In this section we describe the rationale behind the composition of
our model, which consists of five fractions: protein, RNA, metabolites, water
and ions. It should be noted that we did not add lipids and DNA to our
model because we consider only elements that are free to diffuse
through the cytosol. We gathered data from several sources in order to
build a computational model that is representative of the cytoplasm of
{\em Escheria coli}~\cite{Dong1996,Bennett2009,Link1997,Mcguffee2010}.
It has been estimated that a {\em E. coli} cell contains about 50,000 ribosomes per $\mu^3$~\cite{Milo2017a}. The volume of our box is about 12,000 nm$^3$ (\figref{shrink}), therefore in principle one ribosomes molecule should be added, however to reduce the complexity this was omitted.

@@

\begin{figure}[H]
\includegraphics[scale=0.5]{shrinking.pdf}
\caption{A) The initial simulation box with $L=35.0$ nm and the box
  after the shrinking step, $L=22.9$ nm. Biomolecules in green, water in red. B) The evolution of the
  volume of the simulation box during the shrinking simulations. Each
  replica is shown in different colors. }
\label{shrink}
\end{figure}

@@

Magnesium cations were added to our cytoplasm model as counter-ions
for tRNA and ATP. However, as the simulation progressed, FBP and ATP
aggregated around Mg$^{2+}$ and, by consequence, around tRNA
% {\bf,
%   effectively removing the counter ions from the RNA. IS THAT
%   CORRECT?} {\color{blue} No, Figure \ref{fig:tRNA_aggregation} shows that Mg$^{2+}$ is still concentrated around tRNA in the soup }. 
  After identifying this problem, we used simulations of
small systems containing only the metabolites, Mg$^{2+}$ and water to
test different strategies to prevent aggregation in future studies
using long simulations of cytoplasm models. We found that protonating
the phosphate-containing metabolites is enough to reach this goal
(Fig.~\ref{fig:avoiding_aggregation}). Even though ATPH$_3$ and
FBPH$_4$ are not the biologically relevant forms of these molecules in
physiological pH, using their protonated forms may be an acceptable
compromise between parametrization complexity and practical
results. The biggest impact of this approach is on long range
interactions of these metabolites with other molecules due to the
neutralization of their net charge. However, molecules in a crowded
environment are restrained to smaller volumes and shorter inter
molecular distances, which mitigates the effect of charge
neutralization. It has been suggested that charged species should
employ scaled charges in order to be compatible with the empirical
water models~\cite{Leontyev2009a,Leontyev2011a}. Due to the omission
of the electronic polarization, the screening is underestimated and this
corresponds to applying an effective dielectric constant of about 1.7, alternatively scale the charges down by $\sqrt{1.7}$ which would mean that FBP would get a charge of approximately -3 instead of -4.
Due to the high local
density of negative charges in FBP and ATP this might indeed be an
alternative solution, barring use of polarizable
models~\cite{Ponder2010a,Lopes2013a,Ghahremanpour2018b,Walz2018a}.

@@

We submitted this model to a total of \SI{3}{\micro\second} of
molecular dynamics simulations using a modified Amber99sb force
field that was tuned to reproduce experimental data of systems
composed of several macromolecules~\cite{Best2014a}. The model was
validated by comparing to experimental data about the reduction of the
translational diffusion coefficient of proteins in the cytoplasm of
\textit{E. coli} when compared to the dilute condition (Figure
\ref{fig:translational_diffusion})~\cite{Elowitz1999,Konopka2006}. Additionally,
our data is in agreement with recent simulation work and experimental data
that show that the unspecific interactions between macromolecules in a
crowded environment can destabilize
them~\cite{Feig2011,miklos2011,Wang2012b}. Specifically, we can see
that, despite the overall stability of our cytoplasm model even in the
\SI{}{\micro\second} timescale, the proteins are slightly less stable
in the cytoplasm model than in dilute condition (Figures
\ref{fig:structural_integrity_chain} and \ref{fig:rmsf}).
